\section{Introduction}
\label{sec:introduction}

Modern foundation models have shifted the performance frontier of AI inference
from arithmetic throughput to \emph{data coordination}.
When model weights, activations, and key--value (KV) caches exceed accelerator
memory, inference latency is governed less by compute and more by how efficiently
execution coordinates data movement across a hierarchical memory system spanning
GPU Video RAM (VRAM), host Dynamic RAM (DRAM), and Non-Volatile Memory Express (NVMe)
storage~\cite{ma2026inference,gholami2024memorywall}.

This shift reaches industrial informatics first, where inference sits inside
production rather than in a research lab: digital twin synchronization,
predictive maintenance over Industrial Internet of Things (IIoT) sensor streams,
and real-time inspection on the line.
These are closed-loop cyber--physical systems (CPS), and the inference engine
must be not merely fast on average but predictably fast, scan-cycle aligned,
and within the energy envelope whenever the line runs.
Industrial workloads also misbehave on demand---an anomalous sensor triggers
a burst, several models contend for one accelerator, and a
service-level agreement (SLA) cannot be missed.
Bounded response times must align with industrial real-time standards:
Time-Sensitive Networking (TSN) windows and the cyclic Programmable Logic
Controller (PLC) execution semantics of IEC~61131-3.
When inference sits inside the control loop, it must finish within a reserved
sub-window of the scan cycle, or the input--logic--output ordering breaks down.
Latency here is a safety question, not only a quality-of-service one.
The 2\,ms budget used in our evaluation targets bounded
warm-state inference (scoring, short-token continuation), not
autoregressive generation of a 17B model.
Determinism and long-running stability are therefore hard requirements, not
nice-to-haves.

This shift also carries pressing economic and environmental stakes.
Once a model is trained, the recurring cost of \emph{serving} it dominates its
lifetime resource footprint.
Hardware analyses identify memory capacity, bandwidth, and interconnect
latency---not arithmetic throughput---as the binding constraints of
inference~\cite{zhou2024survey}.
The energy penalty of misaligned memory orchestration is substantial: operating
in the I/O-limited regime reduces throughput-per-Watt by more than $3\times$
relative to coordination-aligned operation, raising CO$_2$ output from
15.8\,g to 42.5\,g per $10^6$ tokens at the IEA\,(2024) global carbon
intensity~\cite{iea2024electricity}.
Characterising the \emph{intrinsic limits} of memory orchestration is therefore
a prerequisite for the financial and energy viability of inference-driven AI.

Coordinating memory aggressively is surprisingly easy to do badly:
the same offloading or caching policy can speed up one configuration and
slow down another.
Existing systems reduce memory pressure via CPU/NVMe
offloading~\cite{flexgen2023,aminabadi2022deepspeedinference},
GPU--CPU swapping~\cite{huang2020swapadvisor},
and KV-cache management~\cite{kwon2023pagedattention}.
These techniques deliver substantial gains in some configurations yet show
diminishing returns---or even performance inversion---in others.
Most prior work attacks the problem one direction at a time: model-level work
(quantization, pruning) shrinks the footprint; system-level work (CPU/GPU
offloading, caching) shuffles data between tiers.
Recent papers in IEEE Transactions on Industrial Informatics study resource
allocation, cloud--edge orchestration, and latency-aware scheduling, but reason
at the network layer or about distributed task placement and treat each node's
memory as a black box~\cite{tao2019digital,cao2021survey,chen2021deep,wu2021accuracy,lu2021communication,chekired2018industrial,deng2020dynamical}.
We open that box, modeling how VRAM, DRAM, and NVMe interact inside one serving
node and characterizing cross-tier transfer formally rather than empirically.
This raises a fundamental question: \emph{Is hierarchical memory orchestration
a continuously optimisable problem, or does it admit intrinsic boundaries beyond
which additional coordination cannot improve---and can worsen---end-to-end
latency?}

We answer this question with a regime-based view.
Unlike the Roofline model~\cite{williams2009roofline}, which characterises
compute/memory balance for a fixed workload on static hardware, our framework
characterises a hierarchical memory system under \emph{dynamic} orchestration
where the dominant bottleneck itself shifts as operating conditions change.
Multi-tier inference systems exhibit \emph{phase-like} operational regimes
governed by two dimensionless ratios: the capacity ratio
$\alpha = C_{\mathrm{DRAM}}/C_{\mathrm{VRAM}}$ and the I/O pressure ratio
$\beta = \Delta_{\mathrm{swap}} / [B_{\mathrm{NVMe}}(T_{\mathrm{comp}} + T_{\mathrm{mem}})]$.
Transitions between regimes are \emph{abrupt}: modest changes in model size,
batch size, or bandwidth shift the dominant latency contributor
discontinuously---analogous to a material crossing a phase boundary when a
control parameter passes a critical threshold.
Regime-agnostic orchestration strategies effective within one regime therefore
fail predictably---and for principled reasons---outside analytically
identifiable boundaries.
This regime structure is not LLM-specific; it governs any workload whose active
working set cannot fully reside in fast memory, including vision inference,
retrieval-augmented generation (RAG), and HPC-style scientific workflows.

We make five contributions:
\begin{itemize}
  \item \textbf{Regime-aware analytical model.}
    A multi-tier latency decomposition with explicit boundary conditions that
    distinguish capacity-limited, I/O-limited, and coordination-dominated
    regimes for industrial AI infrastructure, together with closed-form
    prediction formulas for both thresholds: $\theta_C = 0.50$ follows
    analytically from the majority-eviction condition (hardware-independent),
    and $\theta_B = R_B^{\mathrm{sat}} \times (1 + \alpha_{\mathrm{wb}} + \alpha_q)$
    is predictable from three hardware measurements obtainable in a 10-minute
    profiling session---enabling pre-deployment threshold estimation without
    exhaustive regime sweeps.
  \item \textbf{Energy--latency coupling analysis.}
    A direct analytical link between cross-tier transfer dynamics and
    throughput-per-watt under sustained SLA constraints, showing that energy
    per token increases up to $3.2\times$ in I/O-limited regimes.
  \item \textbf{Lightweight multi-tier orchestration.}
    The ORION design and implementation, with Non-Uniform Memory Access
    (NUMA)-aware and swap-aware regime control across VRAM, DRAM, and NVMe,
    comprising three modules: GPU Process Allocator, Task Memory Manager,
    and GPU Booster with swap support.
  \item \textbf{Industrial-scale validation.}
    Long-running workloads (72-hour stress tests) showing better SLA stability,
    energy efficiency, and robustness than strong baselines, with timing
    violation rates meeting IEC~61508 SIL-2 compliance.
  \item \textbf{Workload generality.}
    Regime structure is demonstrated beyond autoregressive LLMs: BLIP2
    visual question-answering, retrieval-augmented generation (RAG), and
    edge YOLOv8 deployments all exhibit the same three-regime behaviour,
    confirming that the framework governs any workload whose working set
    exceeds fast-memory capacity.
\end{itemize}

Together, these results reframe hierarchical memory coordination as
\emph{regime-aware design} rather than continuous tuning, establishing that
hierarchical memory inside an inference node is regime-dependent in a way
that can be written down formally---and once it is, the engineering problem
becomes tractable.
To illustrate the framework's practical accessibility, consider Llama-3\,8B at
batch\,=\,8 on a server with 80\,GB HBM and 16.1\,GB/s PCIe bandwidth:
$W = 41.8$\,GB gives $R_C = 80/41.8 = 1.91$ (fully capacity-compliant,
$R_C \gg \theta_C$).
With nearly all weights resident in HBM, compulsory cross-tier transfer is
small ($D \approx 1.2$\,GB per step); at $\Delta t \approx 120$\,ms,
$R_B = 16.1 \times 0.12 / 1.2 = 1.61$ (well inside the coordination-dominated
regime).
By contrast, constraining HBM to 20\,GB forces $R_C = 0.48$ and
$D \approx 20.9$\,GB of eviction traffic per step, pushing the same workload
deep into the capacity-limited regime (see Supplementary Appendix~A for the
full quantitative derivation).
Any practitioner can compute these two ratios from hardware datasheets and
model-card parameter counts---before running a single experiment.

\FloatBarrier
