\section{Discussion}
\label{sec:discussion}

\subsection{Fundamental limits of hierarchical memory orchestration}

The regime-dependent framework establishes a practically important negative
result: beyond regime boundaries, additional orchestration can predictably
\emph{increase} latency and energy expenditure by exposing synchronisation
overhead without improving throughput.
In capacity-limited regimes, the absence of effective locality imposes
unavoidable memory-access costs that coordination cannot offset.
In I/O-limited regimes, physical bandwidth constraints dominate once cross-tier
transfer approaches sustained limits; any additional coordination overhead is
additive rather than compensatory.
This is analogous to, but mathematically distinct from, thermodynamic phase
transitions.
In a thermodynamic system of diverging size, the order parameter changes
discontinuously and the susceptibility (analogous to $S$) diverges at the
critical point.
Here the ``system size'' is finite: the number of transformer layers $L$ is
fixed, and both $S$ and the boundary width carry $O(1/L)$ corrections from
the continuum limit.
The transitions are therefore more precisely \emph{finite-size crossovers}---abrupt
but bounded changes in the dominant latency term at a well-defined critical
ratio.
Empirically, $S = 4.12$ at $\theta_C$ ($> 2\times S^*$) demonstrates that
the crossover is operationally indistinguishable from a true transition at any
practically relevant model scale, even though $S$ remains finite.
This precision resolves the ``phase-like'' qualifier used throughout the main
text: the transitions are genuine crossovers whose sharpness scales with $L$,
and the analogy to physical phase transitions motivates the dimensionless-ratio
formulation and may extend the framework to other hierarchical-memory systems
beyond AI inference.

\subsection{Reconciling conflicting results in prior systems research}

A regime-based lens provides a unifying explanation for heterogeneous---and
sometimes contradictory---results in prior systems work.
Offloading, caching, and swapping strategies that implicitly assume operation
within coordination-dominated regimes deliver substantial gains when that
assumption holds.
When deployed in configurations where capacity or bandwidth constraints dominate,
their effectiveness diminishes sharply.
Differences in fast-memory capacity, sustained transfer bandwidth, and workload
characteristics can place nominally similar systems in different regimes,
producing divergent benchmark outcomes without obvious explanation.
Making the operating regime explicit therefore directly improves the
interpretability and generalisability of inference systems research.

\subsection{Persistence of regime structure under hardware advances}

A natural counterargument is that emerging memory technologies---HBM capacity
expansion, CXL-based memory pooling, disaggregated memory---will render regime
boundaries irrelevant.
Our analysis shows these technologies can \emph{shift} boundaries by increasing
intermediate capacity or improving sustained bandwidth, but they do not remove
the underlying phase structure.
Once transfer demand approaches a tier's sustained bandwidth ceiling, or once
working-set residency fails in fast memory, the dominant term in the latency
decomposition changes discontinuously regardless of absolute capacity or
bandwidth values.
More sophisticated software strategies---aggressive prefetching, deeper
computation--transfer overlap---can redistribute visible latency within
coordination-dominated regimes but cannot violate physical capacity and
sustained-bandwidth limits.
Beyond those limits, additional coordination becomes additive overhead rather
than a compensatory mechanism.

\subsection{Practical regime-aware inference and limitations}

The framework directly motivates adaptive inference systems.
Rather than applying a fixed orchestration strategy uniformly, a regime-aware
controller can sample the latency decomposition
($T_{\mathrm{mem}},\,T_{\mathrm{swap}},\,T_{\mathrm{sync}}$) at low overhead
and classify the operating regime---a task our depth-3 decision-tree classifier
accomplishes with 93.4\% accuracy across diverse workloads at $<$0.1~ms cost
per classification.
The first question for any memory-bound inference system should be not ``how
much more coordination?'' but ``in which regime does the system operate?''

The regime framework implies concrete prescriptions for each operating
condition.
In the \emph{capacity-limited} regime ($R_C < 0.50$), the primary lever is
fast-memory residency: increase HBM capacity, reduce batch size, or apply
weight quantisation to shrink $W$; additional coordination overhead has
negligible return ($\leq 5\%$ in our experiments, Table~\ref{tab:regime_inversion}).
In the \emph{coordination-dominated} regime ($R_C \geq 0.50$, $R_B \geq 0.40$),
pipelining, KV-cache management, and cross-tier overlap all provide measurable
gains (18--27\%); this is the regime for which existing orchestration tools
are effectively designed.
In the \emph{I/O-limited} regime ($R_B < 0.40$), the correct intervention is
to reduce compulsory transfer volume---via smaller swap granularity, selective
layer offloading, or on-device quantisation---rather than adding scheduling
complexity; aggressive offloading strategies (FlexGen, DeepSpeed-Inference)
can \emph{worsen} latency by 8--12\% in this regime by amplifying an already
saturated transfer path.
These prescriptions are immediately actionable from two ratios computable
before deployment.

This perspective applies beyond large language models to any workload whose
active working set cannot fully reside in fast memory---multimodal inference,
RAG, data-intensive analytics, and HPC-style scientific computing---and
provides a principled foundation for understanding scalability, stability, and
energy efficiency in increasingly complex memory hierarchies.

Several limitations motivate future work.
The regime classifier's 93.4\% accuracy is an in-distribution result:
training and evaluation used measurements from the same five hardware
platforms.
Generalisation to unseen hardware configurations, novel memory-tier
arrangements (e.g., CXL-pooled memory), or architectures with qualitatively
different bandwidth profiles has not been characterised and is an important
direction for future validation.
Large-scale multi-node clusters introduce additional coordination layers that
may shift or fragment regime boundaries; our 4-node experiments observed
$\Delta\theta_B \approx +0.05$ but full multi-node characterisation remains
open.
Hybrid training--inference pipelines and highly irregular sparse-access patterns
may exhibit nested regime structures.
Hysteresis-like latency behaviour observed when workloads traverse boundaries
in both directions suggests \emph{regime memory} driven by nonlinear cache-reuse
effects; characterising this theoretically and exploiting it in scheduling is an
open problem.
The two-ratio framework extends naturally to modern serving paradigms with
variable working sets.
For mixture-of-experts (MoE) models with $N_{\mathrm{exp}}$ experts and
top-$k$ routing, the effective working-set size per step is
$W_{\mathrm{eff}} = W_{\mathrm{param}} \times (k/N_{\mathrm{exp}})
+ W_{\mathrm{act}} + W_{\mathrm{kv}}$, which shifts $R_C$ upward relative
to dense activation and can move an MoE deployment from the I/O-limited to
the coordination-dominated regime at the same hardware operating point.
For speculative decoding, the draft and target models carry distinct working-set
sizes and may occupy different regimes simultaneously, requiring per-model
($R_C$, $R_B$) classification and potentially conflicting optimal orchestration
strategies.
Accounting for these variable-$W$ dynamics is an immediate and natural
extension of the current framework.
Finally, while the regime framework provides first-order prescriptions
(``Practical regime-aware inference'' subsection above), optimal orchestration
strategies \emph{within} each regime---particularly in the coordination-dominated regime where all four
latency terms are co-active---remain open; combining boundary theory with
workload-specific optimisation via reinforcement learning or Bayesian
optimisation is a promising direction.
