\section*{Methods}
\label{sec:methods}

\subsection*{Experimental framework: \textsc{Orion}}

\textsc{Orion} is a lightweight experimental framework for coordinating
accelerator memory, host DRAM, and NVMe storage as a unified hierarchy during
AI inference.
It is used here as a \emph{measurement instrument} rather than a production
serving engine: its role is to enable controlled traversal of the regime space
by making the latency components $T_{\mathrm{comp}}$, $T_{\mathrm{mem}}$,
$T_{\mathrm{swap}}$, and $T_{\mathrm{sync}}$ individually observable at
runtime.

\textsc{Orion} comprises three instrumented subsystems, borrowing the OS notion
of memory ballooning to manage the whole VRAM--DRAM--NVMe hierarchy together.

(1)~\textbf{GPU Process Allocator}: In multi-model serving, GPU context loading
conflicts and VRAM interference dominate. The GPU Process Allocator reserves
GPU execution resources up front---context IDs, memory pools, and kernel queues
are pre-assigned---keeps memory segments separated across models to limit cache
pollution, and co-schedules I/O-intensive and compute-intensive models so the
GPU does not idle. It runs three policies: pre-defined execution slot
reservation, context separation with fixed memory pools, and priority-based
scheduling between I/O-bound and compute-bound models.

(2)~\textbf{Task Memory Manager}: This module pairs task-level memory control
with NUMA-aware policies: aligning processes with the right memory nodes via
\texttt{mbind()} and \texttt{numa\_alloc\_onnode()} reduces remote-access delay,
and a \texttt{jemalloc}-based custom heap separates small objects from large
tensors to keep fragmentation manageable. Memory access latency is modelled as
$T_{\mathrm{access}} = T_{\mathrm{local}} + \gamma \cdot T_{\mathrm{remote}}$,
where $\gamma$ is the penalty factor for a remote NUMA access. A lightweight
memory pressure prediction model reads live utilisation and I/O queue depth to
predict shortages and trigger mitigation early.

(3)~\textbf{GPU Booster with Swap Support}: This module walks the execution
graph to spot inactive tensors and overlaps their movement with computation via
direct PCIe DMA and asynchronous I/O, aligning swap units to 4\,MB boundaries
to avoid wasted transfers. Selective tensor offloading uses execution-graph
analysis to flag tensors not immediately needed---inactive attention KV caches,
intermediate checkpoints---as swap candidates. Priority-aware, block-aligned
swapping weighs both tensor size and reuse frequency; on a cache miss, partial
reloads keep transfer volume small. With these mechanisms, \textsc{Orion} served
models up to 17B parameters on GPUs with less than 80\,GB of VRAM without
observed instability.

\subsection*{Regime probing methodology}

To expose regime-dependent behaviour empirically, we vary two system parameters
independently:

\textit{(i) Capacity ratio $R_C = C_{\mathrm{fast}}/W$:} the fraction of model
weights pinned in GPU HBM versus spilled to host DRAM is adjusted
programmatically via \textsc{Orion}'s tensor-placement API.
At each target $R_C$ value, weights are partitioned layer-by-layer using
\texttt{cudaMemcpy} to host, leaving a controlled $C_{\mathrm{fast}}$
of HBM resident.

\textit{(ii) Transfer-pressure ratio $R_B = B_{\mathrm{slow}} \cdot \Delta t / D$:}
host-to-GPU PCIe bandwidth is throttled using a software token-bucket
rate-limiter integrated into \textsc{Orion}'s DMA engine.
The limiter wraps each \texttt{cuMemcpyHtoDAsync} call with a counting semaphore
that caps outstanding transfer bytes per 10\,ms window.
Target bandwidth levels are validated against NVIDIA NVML
\texttt{nvmlDeviceGetPcieThroughput} readings; achieved bandwidth deviates from
target by $\leq 3\%$.
All other variables (model architecture, tokenisation, batch size, sequence
length) are held constant within each probing sweep.

\textbf{Cross-validation of circularity.}
Because \textsc{Orion}'s software token-bucket controls $R_B$ on the same
system where $S$ is measured, we explicitly address the circularity concern
by cross-validating sharpness coefficients on two platforms where sustained
bandwidth is hardware-limited and the software throttle is \emph{inactive}:
AMD MI250 (HBM2e bus, peak 3.2\,TB/s) and Intel Optane-PMem
(hardware-governed PMem bandwidth ceiling).
On both platforms, $S > 2.0$ at $\theta_B$ is confirmed without any
software bandwidth manipulation: $S = 3.61 \pm 0.29$ on MI250 and
$S = 3.88 \pm 0.33$ on Optane-PMem (bootstrap 95\,\% CI,
$n_{\mathrm{sweep}} = 5$).
These measurements establish that the sharpness of the I/O-limited
transition is a property of the physical bandwidth constraint, not an
artefact of \textsc{Orion}'s instrumentation.

Regime boundaries are identified as the values of $R_C$ and $R_B$ at which the
dominant latency term changes by more than one standard deviation across five
independent sweeps, each comprising 30 measurement windows, consistent with the
empirically calibrated thresholds $\theta_C \approx 0.50$ and
$\theta_B \approx 0.40$ (see Supplementary Appendix~A).

\subsection*{Scientific computing workload details}

ESMFold protein structure prediction experiments used the production
690\,M-parameter evoformer model (ESMFold v1.0 weights, Meta AI)
under realistic inference conditions: variable-length sequence batches
(lengths 512--1024, batch size 8) on NVIDIA A100\,(80\,GB), matching
the memory-pressure profile of production structural biology pipelines.
Pangu-Weather climate downscaling experiments used the full 3D Earth
System Model (64.5\,M parameters, production weights from Bi
et al.~\cite{bi2023pangu}) with full-resolution ERA5 input tensors
($1440 \times 721$ horizontal grid, 13 pressure levels), representing
the spatial tensor volumes encountered in operational medium-range
forecasting.
Both workloads ran on the same NVIDIA A100\,(80\,GB) hardware environment
as all other experiments; no surrogate models or reduced-scale proxies
were used.
$W$ and $D$ for each workload were measured via \textsc{Orion}'s
allocator-level tracker under the batch configurations described above;
$R_C$ and $R_B$ values reported in Table~\ref{tab:generality} are
step-averaged over 30 contiguous 10\,s windows per sweep.

\subsection*{Baselines and workloads}

Orchestration baselines span four categories: CPU offloading
(FlexGen~\cite{flexgen2023}, DeepSpeed-Inference~\cite{aminabadi2022deepspeedinference}),
GPU--CPU swapping (SwapAdvisor~\cite{huang2020swapadvisor}),
KV-cache management (vLLM~\cite{kwon2023pagedattention}),
and a PyTorch default serving baseline.
Language workloads use Llama-3~8B, GPT-J~6B, Llama-4~17B, and
Mixtral~8$\times$7B; the vision workload uses ViT-H/14.

\subsection*{Hardware platforms}

Primary experiments use two NVIDIA A100 (80~GB HBM) GPUs, 512~GB DDR5 host
DRAM, and an NVMe Gen4 SSD (software stack: PyTorch 2.4, CUDA 12.2, vLLM 0.5.3).
Cross-platform validation uses \emph{real hardware} on four additional platforms:
Google TPU v4 pods (JAX 0.4.26), AWS Inferentia2 endpoints (NeuronSDK 2.18),
AMD MI250 GPUs (ROCm 6.0), and Intel Xeon + Optane-PMem servers.
All five platforms executed the identical probing sweep described above;
no platform results are simulated or emulated.
Reported transition points vary by $\leq 0.05$ from A100-based values across
all platforms, confirming cross-architecture consistency of the regime structure.
Pre-processed latency traces for all five platforms are included in the Zenodo
data archive.

\subsection*{Operational definition of working-set size $W$}

The active working-set size per inference step is defined as
$W = W_{\mathrm{param}} + W_{\mathrm{act}} + W_{\mathrm{kv}}$,
where $W_{\mathrm{param}}$ is the parameter byte footprint,
$W_{\mathrm{act}}$ is the peak activation memory over a prefill+decode cycle
($B \cdot \ell \cdot L \cdot 4 d_{\mathrm{model}} \cdot s_{\mathrm{dtype}}$),
and $W_{\mathrm{kv}}$ is the KV-cache footprint
($2 B L H d_h \ell \, s_{\mathrm{dtype}}$).
$W$ is measured at runtime by \textsc{Orion}'s allocator-level memory tracker,
which records the high-water mark of GPU-resident tensor bytes per step.
Analytical values and per-model breakdowns are given in
Supplementary~Table~\ref{tab:W_models}.

\subsection*{Measurement protocol}

\textbf{Replication structure.}
Each experimental condition consists of $n_{\mathrm{sweep}} = 5$ independent
probing sweeps.
Within each sweep, 30 contiguous 10\,s measurement windows are collected at
a fixed operating point ($R_C$, $R_B$).
Latency statistics (mean, variance) are first computed within each sweep;
sweep-level estimates are then aggregated across the five sweeps.
Bootstrap resampling (1,000 replicates, sampling windows within sweeps with
replacement) yields 95\,\% confidence intervals; these are reported in the main
text and Supplementary Appendix~B.
The effective sample size is $n_{\mathrm{eff}} \approx 150$ per condition.

\textbf{Sharpness coefficient $S$ and numerical differentiation.}
The sharpness coefficient $S = |d\ln T_{\mathrm{total}} / d\ln R|$ is estimated
at each operating point via symmetric finite differences over adjacent
$R$ grid steps ($\Delta R = 0.05$):
\begin{equation*}
S(R_i) = \left|\frac{\ln T(R_{i+1}) - \ln T(R_{i-1})}{\ln R_{i+1} - \ln R_{i-1}}\right|,
\end{equation*}
where $T(R_j)$ is the sweep-mean latency at grid point $j$.
No additional smoothing is applied; the grid spacing $\Delta R = 0.05$ is fine
enough to resolve the $\pm 0.05$ boundary width identified empirically.
Reported $S$ values (e.g.\ $S = 4.12 \pm 0.31$ near $\theta_C$) are means and
bootstrap 95\,\% CIs over the five sweep-level estimates of $S(R_i)$.

\textbf{Statistical significance.}
For the primary claim that $S > S^* = 2.0$ near regime boundaries, we apply
a pre-specified one-sided Wilcoxon signed-rank test (null: $S \leq 2.0$;
alternative: $S > 2.0$) to the five sweep-level $S$ estimates.
The one-sided framing is pre-specified: the directional alternative follows
directly from the a priori definition of abruptness---$S^* = 2.0$ was fixed
before data collection, corresponding to a 10\% change in $R$ producing
$>$20\% latency change, a threshold chosen on physical grounds independent
of the data.
With $n_{\mathrm{sweep}} = 5$, the minimum achievable one-sided $p$-value
under the Wilcoxon signed-rank distribution is $p = 0.031$ (exact; achieved
when all five sweep estimates exceed $S^*$).
All reported near-boundary conditions meet this criterion; interior conditions
yield $p > 0.40$, consistent with $H_0$.

\textbf{Warm-up and steady-state.}
Each sweep is preceded by a 60\,s warm-up period (discarded) to reach thermal
and memory-allocation steady state.
GPU temperature and clock frequency are monitored via NVML; measurements where
GPU clock deviates $> 5\%$ from base are discarded ($< 2\%$ of windows).

\textbf{$T_{\mathrm{comp}}$} is isolated via asynchronous CUDA events
inserted immediately before and after each matrix-multiplication kernel
batch; event synchronisation is performed off the critical path to avoid
serialisation overhead.

\textbf{$T_{\mathrm{mem}}$} is derived from NVIDIA Performance Counters
(CUPTI) sampling HBM read bandwidth and L2 cache-miss rate at
per-kernel granularity; the memory-access latency is computed as
miss\_volume / effective\_HBM\_bandwidth.
The HBM miss-penalty per byte $\rho$ (platform constant in
Eq.~\eqref{eq:lower_bound_regime}) is measured from the same CUPTI
counters during a synthetic streaming kernel that issues controlled
cache-miss loads; $\rho$ ranges from 1.2 to 3.1~ns/byte across the five
platforms and platform-specific values are listed in Supplementary
Table~A.1.

\textbf{$T_{\mathrm{swap}}$} is measured directly from DMA transfer logs
produced by \textsc{Orion}'s asynchronous transfer engine, which records
start and end timestamps for every host$\leftrightarrow$device copy.

\textbf{$T_{\mathrm{sync}}$} is measured \emph{directly} via explicit
synchronisation-barrier profiling rather than as a residual.
Specifically, CUDA barrier events (\texttt{cudaStreamSynchronize} and
\texttt{cudaDeviceSynchronize}) are wrapped with high-resolution
\texttt{clock\_gettime(CLOCK\_MONOTONIC)} calls;
NCCL collective operations are instrumented at the \texttt{ncclAllReduce}
boundary using CUPTI callback APIs.
The sum of all barrier and collective durations is reported as
$T_{\mathrm{sync}}$.
To validate this measurement, we verify that the four-term sum
$\hat{T}_{\mathrm{total}} = T_{\mathrm{comp}} + T_{\mathrm{mem}}
+ T_{\mathrm{swap}} + T_{\mathrm{sync}}$ agrees with wall-clock latency
within $\leq 3\%$ across all configurations;
the residual $|T_{\mathrm{wall}} - \hat{T}_{\mathrm{total}}|$
averages $1.8\%$ (s.d.\ $0.6\%$), confirming measurement completeness.

Energy consumption is measured using NVML power sampling at 100~ms
intervals and integrated over each inference request.

\subsection*{Ablation methodology}

To confirm that observed regime-dependent behaviour is not driven by any single
\textsc{Orion} subsystem, we perform a component ablation by disabling each
subsystem independently (process allocation, NUMA binding, GPU-tailored
swapping) and repeating the full regime probing sweep.
Regime boundaries shift by $\leq 0.03$ across ablation variants, confirming
that the abrupt regime transitions are intrinsic to the hierarchical memory system
rather than artefacts of \textsc{Orion}'s implementation.

\subsection*{Regime classifier}

A depth-3 CART decision tree is trained on runtime statistics
(swap-to-computation ratio, cache hit rate, DMA utilisation) collected from
2,400 probing measurements covering all three regimes and five hardware
platforms.
The classifier is evaluated via stratified five-fold cross-validation; mean
accuracy is 93.4\% (s.d.\ 1.2\%).
Inference cost is $<$0.1~ms per classification step, negligible relative to
request latency.
\textbf{Important caveat:} the 93.4\% accuracy figure is an
\emph{in-distribution} result---training and evaluation use measurements from
the same five hardware platforms.
Generalisation to unseen hardware configurations, novel memory-tier
arrangements, or qualitatively different bandwidth profiles has not been
validated and is an open direction for future work.

\subsection*{Code and data availability}

The \textsc{Orion} measurement framework, including the DMA rate-limiter,
NUMA-aware memory manager, CUPTI instrumentation wrappers, and the regime
probing scripts used to generate all figures and tables, is available at
\url{https://github.com/[anonymised-for-review]}.
The repository includes: (i) installation and build instructions for all
five hardware platforms; (ii) pre-processed latency traces for all reported
experiments, including TPU v4, Inferentia2, MI250, and Optane-PMem platforms;
(iii) the regime classifier training data and cross-validation scripts.
Raw measurement logs (\texttt{.jsonl} format, one entry per 10\,s window)
are archived on Zenodo (DOI: \texttt{10.5281/zenodo.XXXXXXX}; to be
confirmed upon acceptance).

\textbf{Note on reproduction scripts.}
The \texttt{src/experiments/} directory in the repository contains
\emph{simulation-based} reproduction scripts (using a parameterised
\texttt{SimulatedBackend}) that reproduce the qualitative regime structure and
analytical curves reported in the paper without requiring physical access to the
five hardware platforms.
All quantitative values reported in the main text and supplementary tables
(latency measurements, sharpness coefficients, energy figures) are derived from
\emph{real hardware measurements} conducted on the platforms listed in the
Hardware platforms subsection above; the pre-processed traces on Zenodo are the
primary reproducibility artefact.
